---
title: 'Homework #5'
author: "Haotian Liang"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

For this assignment (and all assignments in this class), you are to turn in your work as an R Markdown document.  Submit your solutions to GradeScope in PDF format.


# Problem 1

For this problem we are going to use the following dataset of sales of season passes to an amusement park in response to advertising offers.  The dataset contains a yes/no attribute indicating whether or not the recipient of the offer purchased a pass (Pass), an attribute (Promo) indicating whether or not the offer was part of a "bundle" that included parking, and an attribute indicating how the offer was delivered (Channel).

```{r}
season_pass = read.csv("season_pass.csv")
season_pass$Channel = factor(season_pass$Channel, levels=c("Email", "Mail", "Park"))
season_pass$Promo = factor(season_pass$Promo, levels=c("NoBundle", "Bundle"))
season_pass$Pass = factor(season_pass$Pass, levels=c("NoPass", "YesPass"))
head(season_pass)
summary(season_pass)
```

1A) We first want to determine if making the offer as part of a bundle increases the odds that the customer will purchase the pass.  Construct a simple logistic regression model (one predictor) to test this.  How much does including the offer in a bundle increase the odds that it will be purchased?

```{r}
library(tidyverse)
# Change Channel column  text to number 1 represent Email,2 represent Mail, 3 represent Park.
season_pass = season_pass %>% mutate('Channel_1' = case_when(Channel == "Email" ~ 1, 
                                                            Channel == "Mail" ~ 2, 
                                                            Channel == "Park" ~ 3))


# Change Promo column  text to number 0 represent No Bundle,1 represent Bundle
season_pass = season_pass %>% mutate('Promo_1' = case_when(Promo == "NoBundle" ~ 0, 
                                                            Promo == "Bundle" ~ 1))



#Change Pass column  text to number 0 represent No Pass,1 represent Yes Pass.
season_pass = season_pass %>% mutate('Pass_1' = case_when(Pass == "NoPass" ~ 0, 
                                                          Pass == "YesPass" ~ 1))
season_pass <- season_pass[-c(1:3)]
head(season_pass)
```

```{r}
lm.fit <- lm(Pass_1 ~ Promo_1, data = season_pass)

summary(lm.fit)
```
```{r}
Intercept <- 0.4520918
slop <- 0.0968927
with_b_buy <- exp(Intercept+slop*1)/(1+exp(Intercept+slop*1))
with_not_b_buy <- exp(Intercept+slop*0)/(1+exp(Intercept+slop*0))
difference <- with_b_buy-with_not_b_buy
difference
```

  - The odds increase 0.02276349 by offer an bundle.

1B)  Now we would like to determine the effect of the "channel" which the offer is delivered through (email, mail, or in park).  Construct a multiple logistic regression model to test this.  What does this model tell us about the influence of the channel and of including it in a bundle?

```{r}
lm.fit_b <- lm(Pass_1 ~ Promo_1 + Channel_1, data = season_pass)

summary(lm.fit_b)
library(car)
vif(lm.fit_b)
# both Promo_1 and Channel_1 under 5, we can keep them all. 
```

  - From the module we can know there is negative slop of Promo,it means if there is a bundle, it will decrease the possibility of customer to perches the pass. For Channel, there is a positive slop, it means people in park have the most possibility to perches the pass, mail is the second, email is the last.

1C) Is there a statistically significant interaction effect between Promo and Channel?

```{r}
lm.fit_c <- lm(Pass_1 ~ Promo_1 + Channel_1 + Promo_1:Channel_1 , data = season_pass)

summary(lm.fit_c)
```
  - Because of the P value of the Promo_1:Channel_1 is 4.69e-07 is way less than 0.05, we can say there is a significant interaction effect between Promo and Channel.
